\texttt{dispRity} demo for ecologists

### 2015-11-26 doc/dispRity-ecology-demo.Rnw
Ecology demo v2 INIT

\title{dispRity demo for ecological data}

\author{Thomas Guillerme}

@@

@@

\section{Before starting}

\dispRity is a package for calculating disparity in \texttt{R}.
To keep it short, this package allows to summarise ordinated matrices (e.g. MDS, PCA, PCO, PCoA) into single values.
These ordinated matrices can represent the any-o-space of possibilities of your analyses (e.g. for morphological data, the morphospace represents all the possible morphologies).

\subsection{Installation}
You can install this package easily if you use the latest version of \R and \texttt{devtools}.

Note that we use the \texttt{release} branch here which is version 0.1.1.
If you want the piping-hot (and full of bugs) version, change the option \texttt{ref = "release"} to \texttt{ref = "master"}.
This package depends mainly on the \texttt{ape} package and the \texttt{timeSliceTree::paleotree} function.

\subsection{Data}

For this example with ecological data we are going to use data from McClean (unpubl.) that contains an ordination of a distance matrix between different experimental plots.

<<data>>=
## Loading demo and the package data
library(dispRity)
## let's use the matrix from McClean (unpubl.)
data(McClean_data)
## This dataset contains an ordinated matrix (in 20 dimensions) of the distance
## between experimental plots.
ord_matrix <- McClean_data$ordination
## As well as two list of different factors affecting each experimental plot:
## the treatment and the depth.
treatments <- McClean_data$treatment
depth <- McClean_data$depth
@

\section{A classic PCA}
A classical way to represent this ordinated data will be to use PCA plots.

@@

@@

\section{Splitting the data}

@@

>>>>

\title{\texttt{dispRity} demo for ecologists}

%1-Introduce the data
%"the data represents experimental plots in different treatments at two different water depths"

%2-Explain rarefaction better
%"the data represents experimental plots in different treatments at two different water depths"

%3-Hypothesis
%We want to know whether the treatments had an effect on the density of different species in the plots (Does treatment alter invertebrate community composition?). And does depth have an effect on community composition in the experimental plots? Annnnd: Do the treatments affect the communities differently at different depths?

%(Personally I hate them as you know!)  Just use them as a way of visually showing that the 'multivariate' communities in the different treatments/ depths are different or similar and I guess which components explain most of the variance. They really aren't that powerful which is why I love your approach. You sold it so well to me, I think your best selling points (and indeed perhaps actually including a sentence at the start) are that you don't lose all the dimensions of the data like you do in PCA. And that you can keep the important ones in (again like you sold to me, that, some axes may be far more influential than a pca with throw up).

\author{Thomas Guillerme \\ guillert@tcd.ie}

@@

This demo aims to give quick overview of the \dispRity package (v.\Sexpr{packageVersion("dispRity")}) for ecological analysis.
Please refer to \href{https://github.com/TGuillerme/dispRity}{GitHub page: github.com/TGuillerme/dispRity} for other vignettes, namely the \dispRity tutorial that explains the functions in more details.

To keep it short, this package allows to use all the dimensions of ordinated matrices (i.e. PCA, MDS, PCO) for statistical analysis rather than just a sub-set of dimensions.
For example, one might want to know whether the some sort of water treatment alters invertebrate communities and composition in natural habitats.

\tableofcontents

\section{Before starting}
\subsection{Installing \dispRity}
You can install this package easily if you are using the latest version of \R and \texttt{devtools}.

@@

\subsection{Data}
For this example with ecological data we are going to use data from McClean (unpubl.) that contains an ordination of a distance matrix between different experimental plots.

<<data>>=
## Loading demo and the package data
library(dispRity)

## Setting the random seed for repeatability
set.seed(123)

## let's use the matrix from McClean (unpubl.)
data(McClean_data)
## This dataset contains an ordinated matrix (in 20 dimensions) of the distance
## between experimental plots.
ord_matrix <- McClean_data$ordination
## As well as two list of different factors affecting each experimental plot:
## the treatment and the depth.
treatments <- McClean_data$treatment
depth <- McClean_data$depth
@

\section{A classical two dimensional approach}
A classical way to represent this ordinated data will be to use PCA plots.

@@

However, a common problem in this type of multivariate analysis is that often only a handful of dimensions are studied (usually the ones that bears the most variance of the ordinated matrix).

@@

\section{A multidimensional approach with \dispRity}
\subsection{Splitting the data}

@@

\subsection{Testing the hypothesis}

\bibliographystyle{sysbio}
\bibliography{References}
